### 2022-02-25 paper/paper1/tables/ew_table.tex
Break out the EW table into a standalone document

\begin{deluxetable*}{lcccccc}[!ht]
    \tablecaption{Equivalent Width \label{tab:ew_properties}}
    \tablehead{
        \colhead{Star Name} & \colhead{Date} & \colhead{ID} & \colhead{Observed Line Center Position} & \colhead{Intrinsic Line Center Position} & \colhead{Equivalent Width (\AA)}}
    \startdata
    HAT-P-67 & Aug 7, 2020 & & $10328.493\pm0.017$ & & $0.1607\pm0.0059$\\
    HAT-P-67 & Aug 8, 2020 & & $10345.007\pm0.019$ & & $0.1122\pm0.0068$\\
    HAT-P-67 & Aug 8, 2020 & & $10344.993^{+0.015}_{-0.016}$ & & $0.1064\pm0.0051$\\
    \enddata
\end{deluxetable*}