### 2023-05-15 paper/paper1/ms.tex
Will send a draft to Caroline today, reading through now.

\documentclass[twocolumn]{aastex631}
\bibliographystyle{aasjournal}

\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{censor}

\let\pwiflocal=\iffalse \let\pwifjournal=\iffalse
%From: http://arxiv.org/format/1512.00483

\newcommand{\hatp}{\object{HAT-P-67}~}
\newcommand{\hatpb}{\object{HAT-P-67 b}}

%% General
\newcommand{\msini}{\ensuremath{m \sin i}}
\newcommand{\mplsini}{\ensuremath{\mpl\sin i}}
\newcommand{\teff}{\ensuremath{T_{\rm eff}}}
\newcommand{\logg}{\ensuremath{\log{g}}}
\newcommand{\vsini}{\ensuremath{v \sin{I_\star}}}
\newcommand{\feh}{\ensuremath{\rm [Fe/H]}}
\newcommand{\logl}{\ensuremath{\log{L}}}
\newcommand{\vmac}{\ensuremath{v_{\rm mac}}}
\newcommand{\vmic}{\ensuremath{v_{\rm mic}}}
\newcommand{\ms}{\ensuremath{\rm m\,s^{-1}}}
% Activity index R'_HK
\newcommand{\rhk}{\ensuremath{R^{\prime}_{HK}}}
% log of R'_HK
\newcommand{\logrhk}{\ensuremath{\log\rhk}}
% S average value
\newcommand{\Savg}{\ensuremath{\langle S\rangle}}

%% ---------------------------------------------------------------------
%% Solar quantities 
\newcommand{\rsun}{\ensuremath{R_\sun}}
\newcommand{\msun}{\ensuremath{M_\sun}}
\newcommand{\lsun}{\ensuremath{L_\sun}}
\newcommand{\loglsun}{\ensuremath{\log{L_\sun}}}
\newcommand{\teffsun}{\ensuremath{T_{eff,\sun}}}
\newcommand{\rhosun}{\ensuremath{\rho_\sun}}
\newcommand{\loggsun}{\ensuremath{\log{g_{\sun}}}}

%% ---------------------------------------------------------------------
%% Stellar quantities 
\newcommand{\rstar}{\ensuremath{R_\star}}
\newcommand{\mstar}{\ensuremath{M_\star}}
\newcommand{\lstar}{\ensuremath{L_\star}}
\newcommand{\astar}{\ensuremath{a_\star}}
\newcommand{\loglstar}{\ensuremath{\log{L_\star}}}
\newcommand{\teffstar}{\ensuremath{T_{\rm eff\star}}}
\newcommand{\rhostar}{\ensuremath{\rho_\star}}
\newcommand{\loggstar}{\ensuremath{\log{g_{\star}}}}

%% ---------------------------------------------------------------------
%% Earth
\newcommand{\rearth}{\ensuremath{R_\earth}}
\newcommand{\mearth}{\ensuremath{M_\earth}}
\newcommand{\learth}{\ensuremath{L_\earth}}
\newcommand{\teffearth}{\ensuremath{T_{\rm eff,\earth}}}
\newcommand{\rhoearth}{\ensuremath{\rho_\earth}}

%% ---------------------------------------------------------------------
%% Planetary
\newcommand{\rpl}{\ensuremath{R_{p}}}
\newcommand{\mpl}{\ensuremath{M_{p}}}
\newcommand{\kms}{\ensuremath{\rm km\,s^{-1}}}
\newcommand{\lpl}{\ensuremath{L_{p}}}
\newcommand{\teffpl}{\ensuremath{T_{\rm eff,{p}}}}
\newcommand{\rhopl}{\ensuremath{\rho_{p}}}
\newcommand{\ipl}{\ensuremath{i_{p}}}
\newcommand{\epl}{\ensuremath{e_{p}}}
\newcommand{\gpl}{\ensuremath{g_{p}}}
\newcommand{\loggpl}{\ensuremath{\log g_{p}}}

\newcommand{\arstar}{\ensuremath{a/\rstar}}
\newcommand{\zrstar}{\ensuremath{\zeta/\rstar}}

%% ---------------------------------------------------------------------
%% Jupiter
\newcommand{\rjup}{\ensuremath{R_{\rm J}}}
\newcommand{\mjup}{\ensuremath{M_{\rm J}}}
\newcommand{\ljup}{\ensuremath{L_{\rm J}}}
\newcommand{\rhojup}{\ensuremath{\rho_{\rm J}}}
\newcommand{\gjup}{\ensuremath{\g_{\rm J}}}

\newcommand{\rjuplong}{\ensuremath{R_{\rm Jup}}}
\newcommand{\mjuplong}{\ensuremath{M_{\rm Jup}}}
\newcommand{\ljuplong}{\ensuremath{L_{\rm Jup}}}
\newcommand{\teffjuplong}{\ensuremath{T_{eff,{\rm Jup}}}}
\newcommand{\rhojuplong}{\ensuremath{\rho_{\rm Jup}}}
\newcommand{\gjuplong}{\ensuremath{\g_{\rm Jup}}}

\providecommand{\eprint}[1]{\href{http://arxiv.org/abs/#1}{#1}}
\providecommand{\adsurl}[1]{\href{#1}{ADS}}
\newcommand{\project}[1]{\textsl{#1}}

%% ---------------------------------------------------------------------
%% Affiliations
\newcommand{\PSUAA}{Department of Astronomy \& Astrophysics, 525 Davey Laboratory, The Pennsylvania State University, University Park, PA, 16802, USA}
\newcommand{\PSUCEHW}{Center for Exoplanets and Habitable Worlds, 525 Davey Laboratory, The Pennsylvania State University, University Park, PA, 16802, USA}
\newcommand{\Princeton}{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
\newcommand{\RUSSELL}{Henry Norris Russell Fellow}
\newcommand{\CFA}{Center for Astrophysics $\vert$ Harvard $\&$ Smithsonian, 60 Garden Street, MS-16, Cambridge, MA 02138, USA}
\newcommand{\UTAustin}{Department of Astronomy, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
\newcommand{\MCDonald}{McDonald Observatory and Department of Astronomy, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
\newcommand{\UTSpace}{Center for Planetary Systems Habitability, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
\newcommand{\Amsterdam}{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, NL-1098 XH Amsterdam, The Netherlands}
\newcommand{\UCSC}{Department of Astronomy \& Astrophysics, University of California, Santa Cruz, 1156 High St, Santa Cruz, CA 95064, USA}
% -------------------------------------------------------

\begin{document}
\shorttitle{HPF spectral analysis}
\shortauthors{TBD}

\title{A large leading tail of atmospheric escape in the HAT-P-67 b system}

\author[0000-0002-4020-3457]{Michael Gully-Santiago}
\affil{\UTAustin}

\author[0000-0003-2649-2288]{Brendan P. Bowler}
\affiliation{\UTAustin}

\author[0000-0001-9662-3496]{William D. Cochran}
\affil{\MCDonald}
\affil{\UTSpace}

\author[0000-0002-1846-196X]{Aishwarya Ganesh}
\affil{\UTAustin}

\author[0000-0001-9811-568X]{Adam L. Kraus}
\affil{\UTAustin}

\author[0000-0001-9626-0613]{Daniel M. Krolikowski}
\affil{\UTAustin}

\author[0000-0003-2152-9248]{Jessica Luna}
\affil{\UTAustin}

\author[0000-0002-1417-8024]{Morgan MacLeod}
\affiliation{\CFA}

\author[0000-0001-9596-7983]{Suvrath Mahadevan}
\affil{\PSUAA}
\affil{\PSUCEHW}

\author[0000-0002-4404-0456]{Caroline V. Morley}
\affil{\UTAustin}

\author[0000-0001-8720-5612]{Joe P.\ Ninan}
\affil{\PSUAA}
\affil{\PSUCEHW}

\author[0000-0002-9584-6476]{Antonija Oklop{\v{c}}i{\'c}}
\affil{\Amsterdam}

\author[0000-0001-7409-5688]{Guðmundur Stefánsson}
\affil{\Princeton}
\affil{\RUSSELL}

\author[0000-0001-6532-6755]{Quang H. Tran}
\affiliation{\UTAustin}

\author[0000-0002-3726-4881]{Zhoujian Zhang}\thanks{NASA Sagan Fellow}
\affiliation{\UCSC}

% and more...

\collaboration{1}{HPF Helium Exospheres Collaboration}

\begin{abstract}
    To be overhauled...
\end{abstract}

\keywords{Exoplanet atmospheres, Exoplanet evolution, Exoplanet atmospheric dynamics, Stellar winds, Exoplanet atmospheric variability}

\section{Introduction}\label{sec:intro}

Atmospheric escape appears to be a compulsory phase for exoplanets of a certain size and insolation, statistically imprinted in the dearth of 1.5$-$2.0~$R_\oplus$ planets \citep{2017AJ....154..109F}.  The leading candidate physical mechanisms for this rapid transition between mini-Neptunes and super-Earths include photoevaporation \citep{2013ApJ...775..105O,2013ApJ...776....2L,2017ApJ...847...29O} a=nd core-powered mass loss \citep{2019MNRAS.487...24G}.  Whatever the cause, some large fraction of planets must undergo transformative atmospheric escape and the signal should be widely discernable, at least in observationally favorable configurations such as transiting planets.  Such signals have been searched for and increasingly found, with at least 28 detections to date \citep{2022arXiv221116243D}.

Uncertainty in system ages, evaporation timescales, and dominating physical mechanisms, degrade our ability to foretell if any given planet will exhibit ongoing signatures of atmospheric escape.  Episodic stellar wind gusts and other forms of astrophysical variance could also subdue the appearance of atmospheric escape, even where we expect it most.  The 48 published non-detections of atmospheric escape \citep{2022arXiv221116243D} must encode these natural whims in a way that we have not yet disentangled. Nevertheless, we can boost our chances of witnessing active and significant atmospheric escape by targeting sources that seem predisposed to loss.  These intrinsic or extrinsic factors may include proximity to the host star, low surface gravity, and high energy incident radiation.  The small planet radius gap between 1.5$-$2.0 Earth radius range could be expected to harbor atmospheres that have not yet shed their gaseous envelopes, still in the act of eroding to their final rocky state.

Inflated hot Saturns stand out as an especially extreme category that should exhibit mass loss.  Their low gravitational potentials should let go of their atmospheres more readily than their steadfast Hot Jupiter counterparts.  Their lower gravity yields larger atmospheric scale heights making them easier to detect in transmission spectroscopy.  And their large transit depths and short periods should make them easy to detect in large numbers, like Hot Jupiters.

But inflated hot Saturns are rare \citep{2018AJ....155..214T}.  The cause for their underabundance remains an open question, with two leading candidate explanations.  Tidal migration mechanisms may be mass-dependent in such a way as to proceed efficiently for Jupiter-mass planets, but inefficiently for the lower-mass Saturns.  In this scenario, sub-Saturn mass planets never make it to the close-in orbital separations that would lead to the conditions needed for inflation in the first place.

Alternatively---and most consequentially for atmospheric escape---another explanation may prevail.  Inflated hot Saturns may either form \emph{in-situ} or effectively migrate to close-in orbital separations \citep{2018ARA&A..56..175D}, but once they arrive, the intense irradiation overheats the planet.  This heating leads to runaway inflation and ultimately complete disintegration.  In this scenario, the inflationary half-life may be so short as to lead to a low probability of observing members in the class among the overall population, and hence the apparent lack of inflated sub-Saturns \citep{2023ApJ...945L..36T}.

\citet{2011ApJ...738....1B} predicted runaway inflation of hot Saturns as a consequence of the Ohmic dissipation mechanism.  Here, lightly thermally ionized atmospheric flows induce drag in a planetary magnetic field, weakly coupling the stellar incident energy into the planetary interior.  \citet{2018AJ....155..214T} recognized the above two hypotheses for the lack of sub-Saturns, while statistically favoring Ohmic dissipation as the mechanism responsible for inflating Hot Jupiters.  Therefore Ohmic dissipation stands out as a leading physical mechanism for inflation and mass loss.

Recently, \citet{2023ApJ...945L..36T} showed that hot Saturns can undergo catastrophic erosion by stellar X-ray and Extreme UV (XUV) photoevaporative mass loss. Planets with densities less than $\sim$1 g~cm$^{-3}$ achieve mass loss rates of up to $10^3$ $M_\oplus /$Gyr setting up a positive feedback loop: the planet increases in radius, overflowing its Roche lobe, fueling greater mass loss, and increasing in radius even further.  This vicious cycle systematically depopulates the mass-radius plane with a cliff defined by the $\sim$1 g~cm$^{-3}$ dividing line.

A runaway inflation scenario predicts an inevitable and profound mass loss rate for inflated hot Saturns.  As the planet's atmosphere overflows its Roche lobe at an ever-increasing pace instantaneous mass loss rates may exceed $\dot{M}>10^{13}$ g/s, over 10$\times$ larger than those previously seen in atmospheric escape measurements to date \citep{2022arXiv221116243D}.  Inflated hot Saturns therefore make excellent targets for direct measurement of atmospheric escape.

Large mass loss alone does not guarantee detectability.  The ability to detect even immense mass loss hinges on its observability in spectral tracers.  Ly$\alpha$, \ion{He}{1} 10833 \AA, and H$\alpha$ have emerged as the most amenable to detection \citep{2018ApJ...855L..11O,2022arXiv221116243D,2023MNRAS.518.4357O}, but each of these has its own observational limitations.  Ly$\alpha$ can suffer from interstellar medium \ion{H}{1} absorption censoring its low-velocity line core, for example.  Here we focus on \ion{He}{1} 10833 \AA, which offers some advantages.  In particular, metastable Helium's ability to be observed at high spectral resolution from the ground has been especially valuable for evincing velocity substructure of the escaping gas motion relative to the exoplanet restframe \citep{2019A&A...629A.110A,2020ApJ...894...97N}, and convincingly associating the signal to an exoplanetary origin as opposed to stellar contamination \citep{2018AJ....156..189C}.  \ion{He}{1} has resulted in at least 14 systems with detections, including \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N, 2021A&A...647A.129L}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A,2020AJ....159..115K}, HD 209458 b \citep{2019A&A...629A.110A}, WASP 69 b \citep{2020AJ....159..278V}, WASP 52 b \citep{2020AJ....159..278V}, HD 189733 b \citep{2021A&A...647A.129L}, and HAT-P-18 b \citep{2021ApJ...909L..10P}.  An observational picture of mass loss in hot Saturns appears to be lacking, at least in part due to their intrinsic rarity.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/HAT-P-67b_radius_valley.png}
    \caption{Overview of exoplanet demographics.  The pixel bins reflect the observed density of over 5000 planets accessed from the NASA Exoplanet Archive. Detections of atmospheric escape are common among large planets with strong insolation.  \hatpb---among the most inflated planets known---is an outlier.}
    \label{fig:instellation}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{figures/tf2018_fig2_update2023_HAT.png}
    \caption{Mass-radius trends for inflated hot sub-Saturns and hot Jupiters, form modeled after Figure 2 from \citet{2018AJ....155..214T} and updated with NASA Exoplanet Archive confirmed planets.  The trend lines show the mass-radius relationship for equilibrium temperatures of 500, 1000, 1250, 1500, and 2000 K, assuming the mean composition and mean heating efficiency from the original figure.  HAT-P-67~b stands alone in a region defined by the lack of inflated sub-Saturns, and explained by short lifetimes from runaway inflation and mass loss.}
    \label{fig:ThornFortUpdated}
\end{figure}

Here we present a multi-year observational campaign searching for atmospheric escape in \object{HAT-P-67 b}, an inflated hot Saturn orbiting an F type sub-giant at an orbital separation of 0.06 AU and a 4.8-day orbital period \citep{2017AJ....153..211Z}.  Its strong insolation ($T_\mathrm{eq}\sim2000\;$K) combined with \object{HAT-P-67 b}'s extremely low surface gravity makes it an exceptional candidate for strong atmospheric mass loss via Roche lobe overflow.  Its $\sim$0.05 g~cm$^{-3}$ density places it firmly below the 1.0 g~cm$^{-3}$ threshold predicted to exhibit runaway inflation by \citet{2023ApJ...945L..36T}.  Figures \ref{fig:instellation} and \ref{fig:ThornFortUpdated} show how much of an outlier \hatpb~ is: large, low mass, and heavily irradiated.

The evolutionary state of \hatp offers even more intriguing possibilities.  The host star may be an evolved subgiant, opening the prospect for planetary ``re-inflation'' \citep{2022AJ....163...53S,2022AJ....163..120G,2023arXiv230306728G} as the planet's insolation increased with the star's rapid luminosity increase in this part of the short-lived HR diagram.  The tidal gravity of the nearby massive host star may amplify mass loss rates even further \citep{2007A&A...472..329E,2023ApJ...945L..36T}. Its status as a rare inflated Hot Saturn makes \object{HAT-P-67 b} a promising and laboratory, uniquely suited for testing the runaway inflationary predictions of Ohmic dissipation and XUV photoevaporation.

\section{Observations}
\subsection{Habitable Zone Planet Finder (HPF)}

The Habitable Zone Planet Finder Spectrograph \citep[HPF;][]{2012SPIE.8446E..1SM,2014SPIE.9147E..1GM, 2019Optic...6..233M} on the queue-scheduled 10-meter \emph{Hobby-Eberly Telescope} \citep[HET;][]{1998SPIE.3352...34R} operates in the near-IR from $8100-12800~$\AA~ spanning the \textit{z}, \textit{Y}, and \textit{J} bands at spectral resolving power $R=55,000$. The HET fixed-elevation design \citep{2007PASP..119..556S} limits the observability of \object{HAT-P-67} to less than 1 hour ``tracks'' for a fixed range of hour angles before (east track) and after (west track) the star transits the meridian.  Whereas conventional steerable telescopes could conduct continuous point-and-stare observations of \object{HAT-P-67} for hours, HET cannot.  In practice, this limitation means that in-transit and out-of-transit observational phases were rarely possible on the same night.  Instead, we organized the observations into four campaigns to coincide with \object{HAT-P-67 b} transits on the nights beginning 2020 April 27, 2020; May 21, 2020; June 14, 2020; and 2022 April 28.  These campaigns have out-of-transit observations at least one night before and one night after, and often two nights on either side of the transit.  Two more transit snapshots were obtained in 2022 June-July without the accompanying visits immediately before and after.  The in-transit campaigns had up to 14 exposures per HET track, with integration times between 5-8.5 minutes.  We also obtained random-in-phase low-priority ``P4'' queue-filler observations \citep{2007PASP..119..556S}.  These out-of-transit snapshot observations typically received 4 or fewer individual exposures.  We observed HAT-P-67 with HPF for a total of 41 visits on 39 unique nights, with two of those nights observing both the east and west HET tracks.  The total on-source integration time exceeds 13.8 hours, over $N=152$ individual spectra.  Table \ref{tabHPFLog} summarizes the log of observations.

The observation schedule was strategically restricted to the spring season when the Barycentric Earth Radial Velocity (BERV) would Doppler shift telluric absorption lines sufficiently far away from the core of the Helium 10833 \AA~ feature \citep{2022ApJ...939L..11S}.  The small BERV still means the redward Helium line wing has significant telluric contamination between 10834$-$10836 \AA, but the core and blue line wings appear relatively pristine.

Only 19 out of the 39 nights possess A0V telluric calibration standard stars.  These standard stars were used to spot-check our telluric masking.  Figure \ref{fig:TESSoverview} indicates the epochs of select HPF visits overlaid on the \emph{TESS} time-series lightcurve.

\begin{deluxetable}{lccrchr}
    \tablewidth{0pc}
    \tabletypesize{\scriptsize}
    \tablecaption{
        HPF Observation Log
        \label{tabHPFLog}
    }
    \tablehead{
        \colhead{Night-of}   &
        \colhead{Track} &
        \colhead{Desc.} &
        \colhead{BTJD} &
        \colhead{$N_\mathrm{exp}$} &
        \nocolhead{$t_\mathrm{exp}$} &
        \colhead{$\phi$}\\
        \colhead{}   &
        \colhead{} &
        \colhead{} &
        \colhead{(days)} &
        \colhead{} &
        \nocolhead{(s)} &
        \colhead{}
    }
    \startdata
    \input{tables/observation_log.tex}
    \enddata
    \tablecomments{The night-of refers to the local date at the start of the night.  All exposure times were 308.85~s except for observations from 2020-07-21 to 2021-03-30, which were 511.20~s.  Column 2 describes the HET ``track'', restricted to either East (E) or West (W).  The thresholds for the prepositional description in Column 3 is defined in the text.  Column 5 is the normalized phase $\phi \in(-0.5, 0.5)$, with zero at midtransit.}
\end{deluxetable}

\subsection{TESS Light Curves}
HAT-P-67 was observed with the \emph{Transiting Exoplanet Survey Satellite} \citep[TESS,][]{2014SPIE.9143E..20R} in Sectors 24, 26, 51, 52, 53 with 2-minute cadence, and in Sector 25 with 30-minute (FFI) cadence.  The Sector 25 FFI data appeared malformed, possibly due to problems with pipeline CCD smear correction, and was disregarded.

We also assembled a comparison sample of about 1000 lightcurves to interpret the prevalence of lightcurve modulation and stellar activity among broadly F subgiant-like stars.  We selected sources based on \emph{Gaia} DR3 $T_\mathrm{eff}$ estimates and similar $\log{g}$,  and availability of at least one sector of TESS 2-minute cadence data.  We visually spot-checked these lightcurves to understand artifacts and windowing effects.

\begin{figure*}
    \centering
    \includegraphics[width=0.98\linewidth]{figures/TESS_HAT-P-67b_overview.png}
    \caption{Overview of all available TESS Sectors showing 24 full or partial transits with 34 visits with HPF (vertical gray bars), 7 of which coincide with transits.  The rest of the visits sample out-of-transit phases.  The 8 HPF visits between BTJD 2050 and 2690 are not shown.  Note the large time breaks between some consecutive panels, which correspond roughly to the duration of a TESS sector.}
    \label{fig:TESSoverview}
\end{figure*}

\subsection{Gaia DR3}\label{gaiadr3}
The stellar system consists of a binary with an M dwarf companion HAT-P-67B (\emph{Gaia DR3 1358614983131339904}) separated on-sky by $9\farcs0$ \citep{2019MNRAS.490.5088M}.  HAT-P-67A (\emph{Gaia DR3 1358614983131339392}) has a parallax of $2.69\pm0.01$ mas in \emph{Gaia} DR3, placing it at about 372$\pm$1.4 pc.  Minor corrections to the parallax uncertainty \citep{2021MNRAS.506.2269E} and bias \citep{2021A&A...649A...4L} appear negligible for the $G=9.98$ source.  The DR3 parallax places the system about 8.7\% farther than previously estimated by \citet{2017AJ....153..211Z}, which adopted a \emph{Gaia} DR1-informed parallax of $2.92\pm0.23$ mas, including a systematic $-0.325$ mas bias term \citep{2016ApJ...831L...6S}.  The wide companion HAT-P-67B has nearly identical parallax ($2.58\pm0.05$ mas) and proper motions, confirming its interpretation as co-moving, with a projected separation of about 3400 AU.  The IAU naming convention would demand HAT-P-67A\emph{b} to refer to the planet, which orbits the primary.  Hereafter, we simply drop the A designation for notational simplicity, since the wide companion will not factor into our analysis.

\subsection{ASAS-SN, DASCH, and ZTF}
We retrieved ground-based photometry with the \emph{All-Sky Automated Survey for Supernovae} (ASAS-SN) using the Sky-Portal \citep{shappee14,2017PASP..129j4502K}.  The precision of ASAS-SN was too low to perceive stellar variability, so we can place a relatively uninformative limit of $<5\%$ stellar variability on years-long timescales.

Similarly, \hatp appears in the Harvard Plate Archive, with 5809 measurements digitized through the Digital Access to a Sky Century at Harvard (DASCH) program, spanning over 120 years of coarse photometric monitoring.  In principle, these datasets could inform long-term variability trends such as stellar cycles.  In practice, the 0.15 magnitude jitter appears too coarse to perceive any genuine astrophysical variability, with no conspicuous trend seen.  We can therefore place a relatively mild constraint that the star appears stable at the $\sim30\%$ level over periods of tens to hundreds of years.

HAT-P-67A was saturated in ZTF \citep{2019PASP..131a8002B}, but the M-dwarf companion HAT-P-67B had up to thousands of visits across several years.  The data quality appeared too poor to perceive any genuine astrophysical variability, with the indication of some lunar background signals in the periodogram.

\subsection{Other archives}
\authorcomment1{Slated for removal:}  
\textcolor{lightgray}{\hatp has one $R\sim8000$ spectrum from the Intermediate Dispersion Spectrograph on the INT, centered near 4000 \AA.  It has 26 photometry points from \emph{Pan-STARRS} spread across its 5 filters.  The other archival observations have been previously reported by \citet{2017AJ....153..211Z}, including Keck HIRES, Keck NIRC2, and WIYN High-Resolution Infrared Camera.  Overall the system appears relatively unstudied outside of the discovery paper.}

\section{Analysis}

\subsection{Gaia DR3}\label{analysisgaiaDR3}
\citet{2017AJ....153..211Z} previously derived stellar radius estimates of 2.1-2.7 $R_\odot$ through Spectral Energy Distribution (SED) and isochrone fitting as part of a joint orbit fit.  We systematically increase those stellar radius estimates by 8.7\% to match the greater Gaia \emph{DR3} distance (\S \ref{gaiadr3}).  For a fixed $R_p/R_\star$ from the measured transit depth, the larger $R_\star$ implies a proportionally larger planet radius. This update systematically decreases the estimate for the already-low density of HAT-P-67 by 28\%, to a mere $<$0.035 g$\;$cm$^{-3}$, albeit with significant uncertainties from the weak mass constraint.  The luminosity increases by 18\%, to about 10.3 $L_\odot$.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/best_fit_orbit.png}
    \caption{Best fit orbit overlaid on the composite TESS lightcurve.}
    \label{fig:transit}
\end{figure}

\subsection{TESS Light Curve}
Previously, \hatpb~transits had only been detected with \emph{HATNet} \citep{2004PASP..116..266B} and followed up with KeplerCam on the FLWO 1.2 m telescope \citep{2017AJ....153..211Z}. These ground-based photometers were not intended to measure long-term stellar variability signals.  We, therefore, examined the \emph{TESS} lightcurves for out-of-transit photometric variability.  The revised precision and continuous coverage of \emph{TESS} can refine the orbital solution.

\subsubsection{Revised exoplanet orbital parameters}
We assembled a composite lightcurve by stitching TESS Sectors 24, 26, 51, 52, 53, which were reduced with the default SPOC pipeline \citep{2020RNAAS...4..201C}, and lightly post-processed with \texttt{lightkurve} \citep{geert_barentsen_2019_2565212}.  These TESS data exhibit an RMS scatter of better than 1 part-per-thousand (ppt) at native 2-minute sampling.  We fit a Keplerian orbit model to the TESS lightcurve using the \texttt{exoplanet} framework \citep{exoplanet:joss}.  We obtained revised orbital properties shown in Table \ref{tabOrbit}.  These properties are broadly consistent with the previously reported values from \citet{2017AJ....153..211Z} and the updated ephemeris of \citet{2022ApJS..259...62I}.  Figure \ref{fig:TESSoverview} shows an overview of all the TESS Sectors with a green GP trendline in green highlighting the out-of-transit modulations, and the exoplanet transit model in orange.  Figure \ref{fig:transit} shows the best-fit orbit overlaid in purple on the detrended TESS lightcurve, which is binned in the orange dots.  Table \ref{tabOrbit} lists only one of the previous orbit determinations---also assuming a circular orbit---compared with the orbital solution reported here.

\subsubsection{Stellar rotation rate from periodogram analysis}\label{TESSmodulation}

The TESS Sector 26 lightcurve exhibits a weak $\sim$3 ppt peak-to-valley out-of-transit modulation. TESS Sectors 51-53 do not show as conspicuous a modulation signal, but still show some ostensibly stellar variability.  The ebb and flow of the modulation can be seen in the minimally processed TESS lightcurve in Figure \ref{fig:TESSoverview}.  We fit the TESS lightcurve modulation with a quasiperiodic Gaussian Process (GP) model using \texttt{celerite} \citep{celerite1,celerite2}.  The GP model fit to the entire composite lightcurve yields a 4.7-day period; when fit to individual sectors alone the periods hover around 5.9 days.

We can constrain the expected stellar rotation period based on the stellar radius estimate, $v\sin{i}$, and the observation of $i\sim90^\circ$ from orbit fitting and Doppler tomography \citep{2017AJ....153..211Z}.  We adopt a high limit of $v\sin{i}=35.8\pm1.1$ km/s and a lower $v\sin{i}=30.9\pm2$ km/s value if macroturbulence is accounted for.  We obtain a range of $3.2 < P_\mathrm{rot}  < 4.8 $ days.  This range may be typical for F stars that have not yet evolved too far into the sub-giant branch \citep{2022ApJ...930....7A}.  The geometrical constraint comports with the 4.7-day GP-based modulation period derived from the stitched TESS lightcurve, but is lower than the 5.9-day per-Sector fits, slightly preferring the lower 4.7-day modulation as the stellar rotation period.  The 5.9-day value would require a stellar radius over 3.2$R_\odot$, which appears implausible.

\subsubsection{Lightcurve modulation amplitude}
\authorcomment1{Slated for removal:}
\textcolor{lightgray}{We computed the peak-to-valley modulation for the 1000-source \emph{TESS} comparison sample, defined as the difference between the 90$^{th}$ to 10$^{th}$ percentile of each normalized \emph{TESS} lightcurve.  Most lightcurves appear consistent with negligible astrophysical variability and/or variability predominated by instrumental artifacts. \object{HAT-P-67}'s 0.36$\%$ lightcurve modulation ranked in the 87$^{th}$ percentile of the ensemble, about 0.5$\times$ above the median.  Sources more variable than HAT-P-67 exhibited 3-5$\times$ its amplitude of modulation.}

\input{tables/revise_props_tbl.tex}

\begin{figure}
    \centering
    \includegraphics[width=0.98\linewidth]{figures/MIST_only_HATP.png}
    \caption{MIST evolutionary model tracks for HAT-P-67.  The $T_\mathrm{eff}$ and luminosity of \hatp (gray circle) are consistent with either the late main sequence or recently evolved subgiant.}
    \label{fig:evolTracks}
\end{figure}

\subsection{Revised system evolutionary state}\label{secMISTtracks}
The slightly increased distance and radius imply an 18\% more luminous host star than previously estimated.  Figure \ref{fig:evolTracks} shows evolutionary tracks using the solar-metallicity MESA \citep{2011ApJS..192....3P,2013ApJS..208....4P,2015ApJS..220...15P} Isochrones and Stellar Tracks \citep[MIST;][]{2016ApJS..222....8D,2016ApJ...823..102C} for HAT-P-67.  The tracks show that a 1.8 Gyr, 1.64 $M_\odot$ \hatp~ could be at the end of its main sequence lifetime (teal track), yielding a gradual rise in incident radiation over the lifetime of \hatpb.  Alternatively, a 1.54 $M_\odot$ evolutionary track would send \hatp~ along the subgiant branch at an age over 2.2 Gyr, with a rapid 50\% increase in luminosity, potentially leading to ``re-inflation'' of the planet \citep{2021ApJ...909L..16T}.  Either evolutionary track appears consistent with the observed SED, stellar surface gravity, and available constraints from spectral fitting.  We therefore adopt a MIST-based 2$\pm$0.2 Gyr age for the system, in-between the previous Geneva and Dartmouth-based scenarios \citep{2017AJ....153..211Z}.  Additional metallicity and rotation effects could slightly alter the evolutionary tracks and therefore increase uncertainty in inferred ages and masses.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT-P-67b_rv.pdf}
    \caption{RV orbit fit including both HPF and Keck HIRES data points.}
    \label{fig:RVfit}
\end{figure}

\subsection{HPF analysis I: RV fitting}

We reduced and analyzed the HPF data first for orbit fitting via precision radial velocity (PRV) measurements following the procedures described in \citet{2021AJ....161..173T} and updated for joint lightcurve and RV fits \citep{2022AJ....163..225T}.  Here we included the new TESS lightcurves and the existing Keck HIRES points reported in Table 3 of \citet{2017AJ....153..211Z}.  The HPF points exhibited an RV jitter of $\sim164\;\kms$, a few times larger than the typical Keck HIRES measurements.  So even though the HPF points were more numerous, their marginal value for RV orbit determination was subdued owing to the lower RV information content in the near-IR for this relatively rapidly rotating F star.  Figure \ref{fig:RVfit} shows the fitted RV semiamplitude $K=33_{-15}^{+21}$ \kms---assuming a circular orbit---slightly decreasing the uncertainty compared to the previous $1\sigma$ limit.  Table \ref{tabOrbit} lists the revised properties, with minor differences from the existing values.  Notably, our fit does not jointly model the Doppler tomographic signals reported in \citep{2017AJ....153..211Z}, which could further constrain the impact parameter.  Additionally, our input $\rstar$ was slightly lower than the best-fit value derived from \emph{Gaia DR3} and SED fitting.
\authorcomment1{Quang Tran: please verify and revise.}

\subsection{HPF analysis II: \ion{He}{1} 10833 \AA}
The preparation for analyzing the Helium feature followed a similar but slightly different procedure than that used for the RV, namely, we used \texttt{Goldilocks}\footnote{\url{https://github.com/grzeimann/Goldilocks_Documentation}} for 2D echellogram reduction.  This tool outputs 1D extracted spectra for the target, and two reference fibers: blank sky and a laser frequency comb \citep[LFC,][]{2019Optic...6..233M}.  The observations were usually acquired with the LFC turned off, so this unilluminated spectrum was discarded.

The sky fiber and target fiber have slightly different throughputs and illumination properties, with the target fiber receiving $93\%$ of the flux of the sky reference fiber on average.  This ratio depends on wavelength and season at the few percent level.  We quantified the wavelength dependence by acquiring calibration observations of blank sky in both the target and sky fiber.  We applied this wavelength-based scale factor to each target spectrum's associated reference sky fiber to achieve near-imperceptible sky-line subtraction residuals.  A few lines exhibit residuals that may arise from genuine differences in the local atmospheric conditions between the target and sky fiber.

We masked spectral regions predicted to have significant telluric lines with a template generated by \texttt{telfit} \citep{2014AJ....148...53G}.  We shifted the spectral coordinates to their common barycentric-corrected reference frame \citep{2014PASP..126..838W} as implemented in \texttt{astropy} \citep{2013A&A...558A..33A,2018AJ....156..123A,2022ApJ...935..167A}.  We flattened the spectra to two pre-selected continuum indices highlighted as vertical blue bands in Figure \ref{fig:HPFheliumOverview}.  The sky-subtraction, telluric masking, barycentric correction, flattening, and all of our other standard pre-processing steps were carried out in the open-source Python interface \texttt{muler} \citep{2022JOSS....7.4302G}.

Figure \ref{fig:HPFheliumOverview} shows a zoom-in on the He 10830 region of interest, with all 152 individual exposures overlaid.  We see variability in the Helium line of up to 10\%, much greater than the $<1\%$ pixel-to-pixel variation.  The feature width spans about 3~\AA.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_He_spectrum.png}
    \caption{Overlay of all 128 individual HPF exposures of HAT-P-67, spanning 2020-2022. Variability is seen in the \ion{He}{1} 10833 \AA~ triplet, denoted by the vertical orange shaded band, but not in the adjacent regions.  These snapshot spectra were barycentric corrected, and continuum flattened to the regions in the blue vertical bands.  Sharp telluric absorption lines have been masked in regions near 10835 and 10837.5 \AA.}
    \label{fig:HPFheliumOverview}
\end{figure}

The 13.8 hours of exposure, combined with \object{HAT-P-67 b}'s short ($P\sim4.8$ day) period, means that a large fraction of the orbit has been collected, with some phases (especially near transit) heavily sampled, and some other out-of-transit phases sampled more sparsely. For visualization purposes, we constructed a 2-D intensity phase scan, binning in phase and wavelength, and spanning the entire orbit of \object{HAT-P-67 b}.  Figure \ref{fig:HPFscanResid} overlays the planet's approximately $\pm150\;$km/s orbital Doppler velocity, with the stellar rest frame velocity demarcated by the vertical line.  The vanishing $<36$ m/s reflex motion of the star is imperceptible at this scale.  The horizontal dashed lines indicate the moments of transit ingress (-0.03), mid-center (0.0), and egress (+0.03), demarcated as \texttt{TRANSIT} in the figure.

We define 4 additional distinct groups of phases with adjectival qualifiers in anticipation of the need to discuss bulk trends: \texttt{PRE}, \texttt{POST}, \texttt{CONTROL}, and \texttt{BASELINE}.  The \texttt{BASELINE} phases define the out-of-transit baseline.  The ``control'' spectra designate phases just before the baseline, but not used to define the baseline, to inspect minor correlation structure in the spectra without dividing it out.  Table \ref{tabHPFLog} lists the adjectival qualifiers for each HPF visit.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_Helium10830EW_byCampaign.png}
    \caption{Time series EW lightcurve of Helium absorption in HAT-P-67b.  The system exhibits a characteristic absorption leading up to transit, followed by a sharp decline after transit passage.  }
    \label{fig:HPFtimeseries}
\end{figure}

The in-transit phases exhibit over 10\% excess absorption peaks.  A large absorption signal can be seen preceding transit ingress, with some significant absorption evident before $-0.2$ in phase.  We construct a fractional residual absorption spectrum by subtracting off and re-normalizing to the non-varying baseline.  We define the non-varying baseline spectrum as the average over phases $0.3-0.5$, which exhibited stable spectra with the least absorption.  Figure \ref{fig:HPFscanResid} shows that significant absorption can be seen as a few percent at $-0.37$ in phase, rising to 10\% just before and during transit.  The egress drop-off is extremely sharp, with almost no excess absorption directly after transit.

% Describe construction of residual spectra

We estimated bulk characteristics of the Helium excess absorption feature by fitting a Gaussian line profile to the 10828$-$10838 \AA~region of the residual spectra.  The single line initialized at 10833 \AA~ was intended to agglomerate all three components of the metastable triplet into a single feature, consistent with their appearance in Figure \ref{fig:HPFheliumOverview}.  The model constructed in this way has a total of 4 parameters: amplitude $A$, location $\lambda_c$, Gaussian width $\sigma$, and constant offset.  We repeated a similar process for the nearby Si line at 10830 \AA~as a control sample.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/phase_2D_diagram_resid.png}
    \caption{Absorption depth phase scan over the entire HAT-P-67b orbit from 41 visits with HPF.  }
    \label{fig:HPFscanResid}
\end{figure}

Excess is confidently detected from phases $-$0.3 to $+$0.03.  The full-width at half maximum (FWHM) of the feature is about 1.8-2.6~\AA, equivalent to a line-of-sight velocity distribution in the range of 50-75~km/s.  Individual fits to the He feature show a typical line center uncertainty $\pm0.1$~\AA~ or better.  The out-of-transit phases exhibit a line center position of 10833.4~\AA, consistent with the stellar \ion{He}{1} rest frame zero velocity.  The excess absorption resides slightly blueward of the stellar zero restframe velocity reference.  As mentioned previously, the telluric masking near 10836 \AA~ censors our view of the existence (or not) of any redshifted lobe in this region.  A more careful treatment of the telluric absorption could rescue some of the data in this region.  The excess detections exhibit a bulk blueshift of up to 30~km/s.

The excess absorption spectrum exhibits an increasing line-of-sight velocity distribution as the planet transits, going from 60~km/s at ingress, up to 100 km/s at egress.

Finally, we detect a much weaker but still significant post-egress blue-shifted absorption.  This lobe appears to decrease linearly in wavelength, corresponding to a characteristic blueshift increasing from near zero at transit midpoint to 80~km/s at 12 hours after transit.  It exhibits a slightly lower line-of-sight velocity distribution of 50 km/s, about half of the peak during transit.

We examined other spectral lines using the same methodology as above, finding no detectable variability in the \ion{Ca}{2} infrared triplet, Hydrogen lines, or other neutral metal lines.  We show some of the diagnostic plots for these non-detections in Appendix Section \ref{appendixSec}.

\subsection{Keck HIRES}
We retrieved 22 epochs of archival \emph{Keck HIRES} spectra via the Keck Observatory Archive (KOA). Of these, 19 were acquired through the iodine cell as reported by \citet{2017AJ....153..211Z}, providing RV orbit constraints.  The 14\arcsec~ slit height of the C2 decker appears to cause order overlap in \ion{Ca}{2} H and K lines, which contaminated the spectral extraction for 14 of the 22 spectra.  The 3\farcs5 B2 decker allowed the faithful extraction of this region for 6 spectra, which exhibit no perceptible \ion{Ca}{2} H and K variability.  The H$\alpha$ region exhibited no perceptible variability, but the timing of these spectra coincidentally missed the orbital phases ($-0.3$ to $+0.03$) where we see the greatest \ion{He}{1} 10833 \AA~ absorption excess, leaving open the possibility that detectable H$\alpha$ absorption could be present at more favorable orbital phases.  New spectra or a more careful extraction of the archival C2 decker spectra would be needed.

\subsection{Velocity substructure}
Figure \ref{fig:centroids} shows the centroid positions of the Helium excess feature.  The bulk velocity drifts of the planet outflow can be seen with a few conspicuous trends.  The pre-transit phases ($-0.3$ to $+0.03$) show a slight tendency towards blueshift relative to stellar restframe, with an accelerating blueshift from phases $-0.3$ to $-0.15$, albeit with some visit-to-visit scatter.  The post-transit observations ($+0.03$ to $+0.015$) exhibit weak-albeit-significant equivalent widths, as denoted by their smaller marker size.  The centroids dramatically accelerate to larger bulk blueshifts, as low as 10831 \AA.

The densely sampled in-transit phases are shown as a zoomed-in, oversized inset.  Here we can see the planetary rest frame as the dotted line cutting across the panel.  At the moment of planet ingress, the centroids exhibit a redshifted absorption relative to the planet restframe. Measurements from April 2022 and May 2020 transits partially overlap in phase coverage while yielding slightly different line centroid locations.  Both visits appear consistent with near-zero restframe velocity.  The May 2020 visit shows a slight trend towards alignment with the planet restframe.  The April 2020 and June 2020 partial transits sampled phases close to planet egress, with a clear blueshift relative to both the star and planet.

Overall this pattern of velocity centroids appears consistent with the majority of the gas rapidly obeying the stellar restframe, but with a large spread in the line-of-sight velocity distribution as we would expect from a range of launch angles.  We revisit the interpretations and caveats for this substructure in the next sections.

\begin{figure}
    \centering
    \subfloat{\includegraphics[width=5cm]{figures/centroid_overview10833.png} }
    \hskip-3em
    \subfloat{\includegraphics[width=4cm]{figures/centroid_inset10833.png}}
    \caption{Line centroid positions of the  \ion{He}{1} 10830 \AA~ feature.  \emph{left}: The overview panel shows a slight blueshift relative to the stellar restframe, and a weak excess blueshift trend after planet egress. \emph{right:} The detailed zoom-in of the in-transit phase shows a redshift relative to the planet restframe at ingress and a blueshift relative to the planet restframe near egress.  The line-of-sight velocity distribution of 2-3 \AA~ means that the gas distribution exhibits both advancing and receding lobes at all phases other than post-egress.}
    \label{fig:centroids}
\end{figure}

\section{Atmospheric Escape}

\subsection{Size and geometry of outflow}
Helium absorption occurs predominantly before the planet is in-transit, indicating a \emph{leading} tail escaping the planet.  The excess occurs as early as 30\% of the orbital path.  That large fraction of an orbit translates to about 0.1 AU, 130 planetary radii, and well outside of the Roche Lobe, with its Hill radius of merely 2.6 planetary radii.  We are detecting Roche lobe overflow, gas that is entirely unbound from the planet.  The bulk of the detectable unbound gas may be understood as following its own Keplerian trajectory at a slightly shorter period orbit, pulling away from the planet parallel to the orbital path.  Such a scenario may arise from preferential emission on the planet dayside.

The large size of the gas stream overfills the transit chord, which means both the advancing blue side and receding red side of the rotating star are nearly continuously obscured in this spin-orbit-aligned system \citep{2017AJ....153..211Z}. The planet transits near the Stellar equator \citep{2017AJ....153..211Z}, meaning the transit chord covers about 10$\%$ of the projected stellar surface area.  Hypothetically a nearly 100$\%$ opaque optically thick stream of gas covering only the transit chord could reproduce the observed 10$\%$ absorption excess.  Alternatively---and more likely---an optically thin gas stream has some additional vertical extent perpendicular to the plane of orbit, which would act to overfill the entire projected stellar disk.  Assuming the gas has a spatially uniform optical depth of about 0.1 would reproduce the 10\% absorption excess.

\subsection{1D Parker wind models and mass loss rates}\label{pwinds}
The dramatic leading tail of atmospheric escape suggests that more complicated 3D hydrodynamic simulations coupled with stellar wind \citep{2022ApJ...926..226M} may be needed to depict the geometry faithfully.  Here we explore one-dimensional (1D) models, which offer ease of interpretation and rapid calculation.  We employ the open-source \texttt{p-winds} code \citep{2022A&A...659A..62D}, which implements a transonic 1D Parker wind model with radiative transfer of the Helium 10830 triplet \citep{2018ApJ...855L..11O,2020A&A...636A..13L}.

The predicted Helium ionization depends sensitively on the XUV flux \citep{2019ApJ...881..133O}.  Ideally, we would have a measurement of the XUV spectrum of \hatp, but the limited facilities and distance of the source preclude these challenging observations.  Instead, we constructed panchromatic X-ray to visible SEDs by scaling and stitching together synthetic spectra to provide a range of high and low $L_X/L_\mathrm{bol}$.  We chose a 6400 K PHOENIX \citep{husser13} solar metallicity photospheric spectral template with $\log{g}=3.5$, scaled to the solid angle seen by \hatpb.  The F7IV-V \object[tau Boo]{$\tau$ Boo} serves as the closest analog with available synthetic X-ray coronal spectra \citep{2011A&A...532A...6S}.  But \hatp resides closer to the Kraft break than does \object[tau Boo]{$\tau$ Boo}, and the move to hotter F stars may be associated with weakening of the stellar wind and other atmospheric changes leading to the prospect of much lower XUV luminosity for HAT-P-67 \citep{2022ApJ...930....7A}.  So there remains significant uncertainty about the applicability of the \object[tau Boo]{$\tau$ Boo} XUV spectrum to HAT-P-67.  We scaled the synthetic SED of \object[tau Boo]{$\tau$ Boo} retrieved from X-exoplanets \citep{2011A&A...532A...6S} so that the integral of ground-state ionizing photons ($\lambda<504\AA$) exhibited $L_X/L_\mathrm{bol} \in [10^{-7}, 10^{-5}]$.  Figure \ref{fig:XUV} shows two conceivable SEDs with these high and low radiation hardness.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/XUV_flux_schematic.png}
    \caption{Synthetic XUV radiation scenarios for \hatpb.  }
    \label{fig:XUV}
\end{figure}

The SED shows that the XUV corona model does not extend redward to 2600 \AA~photons capable of ionizing out of the 2 $^3$S Helium metastable state.  This apparent deficit should be negligible, since the F spectral type obtains most of its NUV photons ($\lambda\sim2600$ \AA) from the Wein side of the photospheric spectrum, so the hardness of the radiation stems mostly from the assumptions of the corona flux level.  The wavelength-dependent cross section for absorption (not shown) is largest just blueward of the ionization thresholds, so the differences in the spectrum between 504 and 1000 \AA~ have relatively little impact on the overall ionization out of the metastable state.

Equipped with these two SEDs of differing radiation hardness, we explored different mass loss rates and exosphere temperatures.  Figure \ref{fig:pwinds} shows one example model spectrum with $\dot{M} = 10^{13}$ g/s, and $T_0=9100$~K, with the a medium XUV spectrum ($L_x/L_\mathrm{bol}=10^{-5}$).  This mass loss rate would imply a characteristic lifetime of $\sim$2000 Myr.  The model spectrum exhibits approximately the same equivalent width as our HPF observations, making it a hypothetical scenario among a family of partially degnerate solutions.

At least a few shortcomings limit the applicability of this 1D Parker wind model.  First, the model-dependent line profile (\emph{i.e. width and depth}) is degenerate with XUV flux, $\dot{M}$, and $T_0$, and so a range of these parameters can be fine-tuned to obtain a large range of mass loss rates consistent with the data and our limited understanding of the XUV flux.  These known degeneracies have been pointed out previously \citep{2022AJ....164..234V,2019ApJ...881..133O}, but the problem for \hatpb~ appears somewhat more acute since XUV data are scarce for F-type stars near the Kraft break.  Second, the 1D model breaks down when attempting to explain the inherently 3D leading tail geometry.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/pwinds_data_overlay.png}
    \caption{Simulated 1D Parker wind model of Helium absorption in \hatpb.  The spectrum was generated with a mass loss rate of $10^{13}$ g/s, exosphere gas temperature of 9100 K, and with an XUV luminosity $L_x/L_\mathrm{bol}=10^{-5}$ intermediate between those shown in Figure \ref{fig:XUV}.  The lines have about the same  equivalent width, while additional unmodeled velocity dispersion of the gas can explain differences between the line shapes of model and data.  Overall the 1D models may be too simplistic to represent the inherently 3D structure of the outflow.}
    \label{fig:pwinds}
\end{figure}

\section{Causal Interpretations} \label{secDiscuss}

\subsection{Causes for the leading tail} \label{secLeading}
Several physical phenomena could give rise to a leading tail geometry.  Here we explore the predominant two physical effects: orbital shear and stellar wind confinement.

Figure \ref{fig:KeplerianShear} shows an illustration of Keplerian shear adapted to the system properties of \hatpb.  In this shear-dominated scenario, the planetary wind launches primarily from the dayside, with relatively little or no wind launched from the nightside.  The planet wind initially launches radially outward from the exobase, with the strongest wind located near the sub-stellar point, the line connecting the planet to the star along the vertical axis in the figure.  Inefficient grazing incidence heating near the terminator subdues the mass loss in the $x-$direction, meaning that this wind exhibits not a hemispherical shape, but more concentration along the star-planet line.  The gas increasingly experiences the star's Keplerian potential past the Roche lobe, accelerating in the direction of orbital motion, $+x$.  The accelerating column eventually overtakes the planet completely, with the prospect of extending to hundreds of planetary radii.  The mass loss has to be large enough to shield the material to make it observable in the metastable \ion{He}{1} 10833 \AA~triplet.

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/KeplerianShear_v0p3.png}
    \caption{Schematic of Keplerian orbital shear.  Locations outside the Roche lobe experience orbital shear from the Keplerian potential.  A wind launched primarily from the dayside will tend to form a leading tail.}
    \label{fig:KeplerianShear}
\end{figure}

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/homogeneous.png}
    \caption{Isotropic mass loss}
    \label{fig:IsotropicMdot}
\end{figure}

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/asym.png}
    \caption{Asymmetric dayside mass loss}
    \label{fig:DaysideMdot}
\end{figure}

Stellar wind confinement \citep{2022ApJ...926..226M} arises from a bow-shock configuration, in which the planetary wind encounters the stellar wind and a shock front causes a density enhancement.  A trailing tail may also be present, but would not undergo the same boost in density, and would be rapidly carried away alongside the stellar wind.  This asymmetry could yield the appearance of a leading tail even if mass loss proceeds isotropically, or the asymmetry could be reinforced by an additional dayside/nightside asymmetry.  Examples of this asymmetry can be seen in 3D MHD modeling of WASP 107~b \citep{2022ApJ...926..226M}.

The overall observation of a leading tail comports with either of these two physical mechanisms, and we suspect both are at play in \hatpb.  Both cases appear needed to create the conditions of high enough column density to populate the \ion{He}{1} metastable triplet to detectable levels while imbuing the velocity substructure that we see.  Here we interpret the bulk velocity substructure in Figure \ref{fig:centroids} under a combination of the two.  We attribute the pre-transit arc to stellar wind acceleration.  The stellar wind is initially too weak to plow the dense escaping gas until a separation of 60-130 planetary radii ($+0.15$ orbital phase), the inflection point in an enormous bow shock.  Gas at these separations has had enough time to diffuse, both out of the orbital plane and to slightly larger radii shells, in turn lowering the column density along the line-of-sight.  The lower column density provides both fewer Helium atoms to participate in absorption, and less overall NUV shielding, leading to greater fractional ionization and further depopulation of the \ion{He}{1} metastable state.

The weak post-egress tail can be understood as follows.  A weak nightside mass loss means that there is both less overall column density and less NUV photoionization shielding, subduing the overall \ion{He}{1} metastable state's signal strength.  The blueshift arises from the Keplerian shear that lags the planet, and the ever-blueshifting centroid represents the stellar wind's greater ability to carry away the lower total inertia of less nightside material.

In order to examine the in-transit velocity structure, we have to consider the interplay of a two subtle geometrical effects.  First, the planet's Roche lobe is small enough (2.7 $R_\mathrm{p}$) that the majority of the projected stellar disk should be filled with gas unbound from the planet, even at the time of mid-transit.  In other words, only a small fraction of the \ion{He}{1} excess signal would be expected to trace the planet's motion, as opposed to the hypothetical in-transit signal for WASP-107~b that definitively traces out the planet's orbital path \citep{2022ApJ...926..226M}.

Second, the star's non-negligible rotational velocity sets up a configuration analogous to Doppler Tomography in the Rossiter-McLaughlin (RM) effect, but distinct in a subtle way.  Whereas traditional Doppler Tomography treats the scanning reticle as an opaque planetary disk, here the reticle resembles a transmissive filter with a wavelength center and width that varies in space and time.  We rank this effect as second-order and return to it in the Discussion Section \ref{vsini-interplay}.

\subsection{Ohmic Dissipation and runaway escape}
\object{HAT-P-67 b} exhibits both dramatic atmospheric escape and significant radius inflation.  Ohmic dissipation stands out as a physical mechanism that could be responsible for both of these outstanding attributes.  \citet{2011ApJ...738....1B} showed that Ohmic heating acts to inflate hot Jupiters, with planets $\lesssim 0.5 M_\mathrm{Jup}$ overflowing their Roche lobes and leading to evaporation on Gyr timescales.  The Ohmic heating scenario requires only a modest planetary magnetic field ($\gtrsim$1 G) and an equilibrium temperature great enough to thermally ionize some modest fraction of neutral metals, such as the low ionization species of \ion{Na}{1} and \ion{K}{1}.  The ``sweet spot'' for this phenomenon appears to prefer equilibrium temperatures in the range of $1500<T_\mathrm{eq}<2000$ K \citep{2011ApJ...738....1B,2012ApJ...745..138M,2016ApJ...819..116G,2018AJ....155..214T,2022A&A...658L...7K}, where the conductivity is strong enough to cause an effective drag without being so strong that magnetic braking slows the planetary wind.  HAT-P-67 b's $\sim$2000 K sits to the higher end, but still in a region of high Ohmic dissipation heating efficiency.  A proposed order-of-magnitude scaling law predicts an inflation timescale, Eq. 20 in \citet{2011ApJ...738....1B}:

\begin{equation}
    \tau_{\textit{infl}}\sim \left(\frac{0.01}{\epsilon} \right) \left(\frac{M}{M_{J}}\right)^2 \left(\frac{R_{J}}{R}\right)^3 \left(\frac{1500 \textrm{K}}{T_\mathrm{eff}}\right)^4 \textrm{Gyr}. \label{eqInflate}
\end{equation}

\noindent yielding an incredibly short $<$5 Myr timescale for \hatpb, assuming a typical $\epsilon\sim0.01$.  The order of magnitude of this inflation timescale is so fleetingly short that---according to this scenario---\hatpb~must be in the runaway stage of inflation, rapidly losing mass, and growing in surface area to fuel a positive feedback loop.  Under this interpretation, \hatpb~ would represent an example of a new category of planet system that is doomed to evaporate entirely due to the Ohmic Dissipation mechanism, as predicted by \citet{2011ApJ...738....1B}.  At Roche lobe overflow, a 5 Myr inflation timescale may imply an instantaneous $\dot{M}_\mathrm{infl}>10^{3}$ M$_\oplus$/Gyr.

A few caveats complicate the unambiguous causal interpretation of Ohmic dissipation.  Chiefly, numerical simulations \citep{2013ApJ...763...13W} and analytic theory \citep{2016ApJ...819..116G} indicate that Ohmic dissipation can stall the contraction of hot Jupiters, but cannot easily ``re-inflate'' them after having undergone a traditional cooling curve.  Heat transfer from the atmosphere to a cooled core simply proceeds too slowly, on order of tens of Gyr \citep{2016ApJ...819..116G}.  This path dependence to Ohmic dissipation would restrict the allowed evolutionary histories of \hatpb, to have arrived to its current location within a few to tens of Myr.  This short time window prefers a physical mechanism such as disk migration, which would be faster than secular eccentric migration with an outer companion. \emph{In-situ} formation of a hot Saturn at these close-in separations may be implausible \citep{2018ARA&A..56..175D}.  Other caveats like unknowns in planetary magnetic fields, zonal band geometries, and dayside/nightside temperature differences make it impossible to uniqely prescribe Ohmic dissipation, and instead several additional factors may also be at play \citep{2021A&A...645A..79S}.

\subsection{XUV Photoevaporation}
XUV photoevaporation heats the upper layers in the atmosphere, driving a hydrodynamic wind \citep{2009ApJ...693...23M}.  Photoevaporation stands out as offering a natural cause of the observed leading tail attributable to dayside/nightside differences in mass loss: the greatest supersonic motions arise from the sub-solar point on the planetary dayside.  The effect may be boosted if dayside/nightside energy transport proceeds inefficiently \citep{2009ApJ...693...23M}.  Such a scenario is illustrated in Figure \ref{fig:KeplerianShear}.

The combined effects of photoevaporation, anomalous heating, and tidal gravity are predicted to have especially drastic outcomes for inflated hot Saturns \citep{2023ApJ...945L..36T} such as \hatp, where the low densities lead to large mass loss rates:

\begin{equation}
    \dot{M} = \frac{3}{4}\frac{\eta F_\mathrm{XUV}}{G K_t \rho_\mathrm{XUV}}\label{thorn23Mdot}.
\end{equation}

Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ greater than at present, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^-3$ threshold, at which point the mass loss rate increases to its current value.  Having lost a significant portion of its mass, its orbit expands to conserve angular momentum.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

\subsection{Constraining mass loss rate}
The theory-inspired mass loss rates above correspond to 1D Parker wind \ion{He}{1} feature depths and widths similar to what we observe, as evinced in Figure \ref{fig:pwinds}, with $\dot{M}_\mathrm{1D}\sim100$ M$_\oplus$/Gyr.  %\authorcomment1{To be continued--- interpret this congruence}.

First, the scaling law arguments that produced Equation \ref{eqInflate} were only proposed as order of magnitude estimates not to be taken too faithfully.  So a $10\times$ higher inflation timescale of 50 Myr would yield $\dot{M}_\mathrm{infl}\sim10^{14}$ g/s, still a very large mass loss rate, but only a factor of 5 away from the baseline 1D model.  Factors of a few uncertainties in the Ohmic dissipation efficiency $\epsilon$ may also contribute.

As mentioned in Section \ref{pwinds}, inferring the mass loss rate from observations is strongly model dependent.  The ionization fraction profile, exosphere temperature $T_0$, and $\dot{M}$ are all at least partially degenerate.  In particular, the uncertain XUV-to-NUV flux ratio confounds the interpretation of the observed Helium signal.  Even the ``high XUV'' scenario in Figure \ref{fig:XUV} does not have enough ground-state ionizing photons to penetrate to within 5 $R_p$ of the planet's surface, in a 1D Parker wind model.  This means that at best only 1 in $10^{6}$ Helium atoms participate in the observability of the Helium metastable triplet lines in the baseline 1D Parker wind model.  Turning down the XUV flux makes the signal even less observable.  A larger mass loss rate would be needed to explain the same observed HPF Helium absorption, if say only 1 in $10^{7}$ helium atoms are populating the metastable state.

\subsection{Reinflation}
Evolved stars increase in luminosity, delivering greater insolation to planets at a fixed separation.  The heightened equilibrium temperature can cause mature planets to inflate, the phenomenon known as \emph{reinflation}.  Such re-inflated hot planets around red giant stars have been recently found by \emph{TESS} \citep{2022AJ....163..120G,2023arXiv230306728G}, though not all planets around evolved stars appear to re-inflate \citep{2022AJ....163...53S}.

\citet{2017AJ....153..211Z} estimated that \hatpb~received about twice the incident flux as a Zero Age Main Sequence (ZAMS) \hatp, based on comparison to the Geneva isochrones. In Section \ref{secMISTtracks} MIST evolutionary tracks showed two equally plausible states: the tail-end of the main sequence or a recently evolved sub-giant.  Figure \ref{fig:OhmicInflate} employs these tracks to assess the re-inflationary impact of \hatpb~could be, assuming that the orbital location has not changed over the system lifetime.  The second panel from the top illustrates a countervailing effect: the decrease in Ohmic dissipation heating efficiency, $\epsilon$, could counteract the increase of stellar flux, assuming the functional form for $\hat{\epsilon}_G$ described by \citet{2018AJ....155..214T}.  Together the effects nearly cancel when computing the inflationary lifetimes, yielding nearly identical curves in the bottom panel.  The third panel recreates the numerically computed curve for $0.5\;M_\mathrm{Jup}$, $T_\mathrm{eq}=$1800 K from \citep{2011ApJ...738....1B}, which should be considered as representative since it was not necessarily tailored to the evolutionary history of \hatpb.  Nevertheless, the similarity of the curve is remarkable since it arrives at approximately the correct planet radius at the right age.  The modeled mass is slightly higher than, but consistent with the RV-measured mass, and would imply a large and increasing mass loss rate over its lifetime.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/reinflation_MIST.png}
    \caption{Conceivable evolutionary scenarios for \hatpb.  The planet's equilibrium temperature increases as the stellar luminosity gradually increases on the main sequence.  The Ohmic dissipation model makes predictions for runaway inflation, depending on the incident stellar flux.  The heuristic scaling relation for the inflation timescale predicts unusually small values in this extreme regime but illustrates the countervailing effects of the top two panels.}
    \label{fig:OhmicInflate}
\end{figure}

A doubling of insolation would place the young \hatpb~at a 20\% lower equilibrium temperature of 1600 K.  Perversely, the $\epsilon_G(T_\mathrm{eq})$ may tend to \emph{decrease} over time since the heating efficiency appears to peak at 2.5\% at 1600 K \citep{2018AJ....155..214T}, and reduce to about 1\% at 1900 K.  So it's possible that the inflation timescale would remain relatively unchanged over the planet lifetime, despite a doubling of incident radiation.

\subsection{Evaporative annihilation as the cause of a lack of sub-Saturns}
The Ohmic dissipation mechanism made two key observable predictions.  Heating efficiency should initially increase as a function of equilibrium temperature, then degrade as irradiation exceeds equilibrium temperatures of about 2000 K.  Second, hot Saturns ($\lesssim0.5 M_\mathrm{Jup}$) should undergo runaway evaporation, whereas hot Jupiters ($\gtrsim 1 M_\mathrm{Jup}$) should reach stable equilibrium---albeit inflated---radii after Gyr timescales.  Both of these outcomes predated data that could adequately validate them, and these outcomes differ from the behavior of other heating mechanisms, such as photoevaporation or tides.  When applied to an ensemble of planet systems, the Ohmic dissipation mechanism would expect a dearth of inflated hot Saturns.

\citet{2018AJ....155..214T} statistically favored Ohmic dissipation based on evidence for a turnover in the heating efficiency of hot Jupiters in a sample of 281 Giant planets.  Furthermore, \citet{2018AJ....155..214T} also highlighted the lack of inflated sub-Saturns as an anomaly: by that time, the observational capability existed to detect such sources in large numbers and yet they were not found.

This dearth of hot Saturns has persisted into the \emph{TESS} era.  Figure \ref{fig:ThornFortUpdated} highlights the region of highly irradiated close-in planets with $N>5000$ planets accessed through the NASA Exoplanet Archive.  \hatpb~ occupies a relative void in this region.

The observation that \hatpb---the most inflated hot Saturn searched for atmospheric escape to date---exhibits mass loss indicative of a terminal fate leads us to a natural explanation for the dearth of inflated hot Saturns.  The inflationary lifetime of hot Saturns is so short that observing planets of this class is unlikely.  \hatpb~ appears to be one of the rare planets that we happen to catch a glimpse of before it has been annihilated.  The lack of inflated hot Saturns is analogous to the small planet radius gap \citep{2017AJ....154..109F}, with Ohmic dissipation replacing core-powered mass loss as a heating mechanism, and with photoevaporation still at play.

We, therefore, attribute this cliff in the $M_p$ vs $R_p$ diagram to either Ohmic dissipation, XUV photoevaporation, or both.  \citet{2018AJ....155..214T} emphasized the positive feedback loop as a cause for the lack of sub-Saturns and noted an alternative interpretation-- mass-dependent migration segregation.  While both scenarios are conceivable, the evidence of strong evaporation of \hatpb~ favors the evaporation scenario.  Furthermore, the runaway evaporation of hot Saturns was naturally predicted by the Ohmic dissipation mechanism, whereas mass-dependent migration has only been developed as a post-diction.

One key observation may tip the scales in favor of XUV photoevaporation as the predominant cause of mass-loss.  The conspicuous leading tail favors preferential mass loss on the highly irradiated planet dayside.  While Ohmic dissipation would exhibit preferential dayside heating, it is the ability of XUV photons to create a planetary wind that tilts the balance.

Finally we consider the prospect that the planetary mass is much lower than its 1-$\sigma$ bound, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This extreme scenario could manifest enormous mass loss rates, with the planet atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.  This terminal Roche Lobe overflow would have several observational predictions.  Heavy elements woud easily leak into the planetary wind, yielding possibly many observable metal lines in the UV.  These may be observable even under more mild conditions, due to lofting of metals into the upper atmosphere, so some care would be required to quantitatively distinguish these two scenarios.

\begin{figure}[t]
    \includegraphics[width=\linewidth]{figures/tf2018_fig2_theory.png}
    \caption{Ohmic dissipation runaway inflation timescale in the exoplanet mass-radius diagram.  Individual planets with confident mass and radius detections are shown as small black dots. Helium non-detections are shown in red squares with all detections of atmospheric escape coming from \citet{2022arXiv221116243D}.  \hatpb~ resides in a sparsely populated region with a vanishingly short inflationary timescale, making it unstable to runaway evaporation with large mass loss expected.}
    \label{fig:tf2018_theory}
\end{figure}

As seen in Figure \ref{fig:ThornFortUpdated}, \hatpb~has few other known analogs.  Figure \ref{fig:tf2018_theory} illustrates why: the inflationary lifetime for planets with comparable mass, radius, and $T_\mathrm{eq}$ is exceptionally short, and so they are rare.  We constructed the $\tau_\mathrm{infl}$ contours in the following way.  We censored the $\tau_\mathrm{infl}$ to regions in the mass-radius plane where numerical simulations of Ohmic dissipation resulted in runaway inflation \citet{2011ApJ...738....1B}.  We interpolate the $T_\mathrm{eq}$ trendlines from Fig. \ref{fig:ThornFortUpdated} to compute $\tau_\mathrm{infl}$ from Equation \ref{eqInflate}, including the best fit Gaussian prescription for $\hat{\epsilon}_G$ from Eq. 34 of \citet{2018AJ....155..214T}.  This trend should exhibit significant uncertainty owing to the order-of-magnitude nature of the scaling relation, source-to-source scatter in $\hat{\epsilon}_G$, and to what extent $\hat{\epsilon}_G$ applies for planets less than $0.5 M_\mathrm{Jup}$, which were excluded in the derivation of $\hat{\epsilon}_G$ owing to their scarcity.  Nevertheless, this contour appears to match the bulk trends: planets tend to reside in either the non-inflationary, stable locus or the inflationary-yet-stable locus.  A relatively few $\sim$dozens of planets approach the ``cosmic shoreline'' of inflationary lifetimes in the tens of Myr. The detection of atmospheric escape along this frontier (green circles in Figure \ref{fig:tf2018_theory}) supports this interpretation.  More observations and theoretical modeling of atmospheric escape along this ridge could help better understand its realized versus expected shape.

\subsection{HAT-P-67 b analogues}

The key figure-of-merit can be distilled to $\tau_\mathrm{infl}$, which can be thought of as an atmospheric escape spectroscopy metric \citep{2018PASP..130k4401K} for inflated planets in the shaded regime of Figure \ref{fig:tf2018_theory}.  Table \ref{tabHeSM} presents the rank-ordered list of planets by this metric, indicating that they are mostly smaller and more massive than \hatpb.  Only KELT-11~b shows a shorter nominal inflationary timescale than \hatpb, owing to its low mass and equilibrium temperature residing closer to the peak of $\epsilon_G(T_\mathrm{eq})$.  We predict that these few other inflated \hatpb-analogs should show evidence for significant atmospheric escape, comparable to what we see for \hatpb.  These sources make excellent targets for \ion{He}{1} 10830 observations, and/or other atmospheric escape diagnostics.  As emphasized, there are significant uncertainties in the inflation timescale, but at least its order-of-magnitude value gives us a quantitative and justified way to prioritize target selection.

\begin{deluxetable}{lLLCR}
    \tablewidth{0pc}
    \tabletypesize{\scriptsize}
    \tablecaption{
        Coarse Inflation Timescales for Hot Jupiters
        \label{tabHeSM}
    }
    \tablehead{
        \colhead{Planet}   &
        \colhead{Mass} &
        \colhead{Radius} &
        \colhead{$T_\mathrm{eq}$} &
        \colhead{$\tau_\mathrm{infl}$} \\
        \colhead{}   &
        \colhead{$M_\mathrm{Jup}$} &
        \colhead{$R_\mathrm{Jup}$}   &
        \colhead{K} &
        \colhead{Myr}
    }
    \startdata
    \input{tables/tab_HeSM.tex}
    \enddata
    %\tablecomments{Inflation timescale assumes $\epsilon=0.01$.}  Not true, we use the empirical estimate
\end{deluxetable}

\section{Discussion}
%\subsection{Prospects for a non-circular orbit}
%\authorcomment1{Slated for removal:}  
%\textcolor{lightgray}{We premised our joint \emph{TESS} lightcurve and HPF RV time series orbit fit on a circular orbit. Here we consider how a finite eccentricity orbit could effect the observations and lead to the illusion of a leading tail... [...]}

\subsection{Cause for stellar modulation}
As discussed in Section \ref{TESSmodulation}, the \emph{TESS} lightcurve exhibits up to $0.36\%$ peak-to-valley modulation amplitude, with a characteristic timescale of $P=4.7-5.9$ days. Some sectors exhibit lower amplitudes.  The conventional interpretation would consign these cyclical modulations to the familiar stellar activity: surface features---either starspots, faculae, or plage---entering and exiting the projected stellar disk on the stellar rotation period comparable to the 4.8-day orbital period of planet b.  The cause for such orbital and rotational synchronization would then be either coincidence or secular star-planet tidal interactions.  For the latter, disk migration could have naturally ceased near the co-rotation radius, leaving the planet to reside naturally near the stellar $P_\mathrm{rot}$.

Alternatively, Star Planet Magnetic Interaction (SPMI) could be at play.  In this scenario, the magnetic field of the planet permeates the space between the star and the planet.  Magnetic perturbations propagate via planetary magnetic fields with the Alfv\'en speed.  The planet can interact with the star if the Alfv\'en speed exceeds the stellar wind speed controlling the bulk motion of the intervening medium.  The non-detection of variability in the \ion{Ca}{2} H and K lines and H$\alpha$ lines appears to disfavor this SPMI interpretation, since SPMI could be expected to cause variations in these diagnostics.

It is hypothetically possible---albeit unlikely---that mass loss could be great enough to produce variability in the wide band \emph{TESS} lightcurve.  The HPF spectra reveal up to 10\% signal depth over a few Angstroms.  The TESS bandpass barely includes 10830 \AA, at a location of diminishing throughput, so it appears impractical that Helium alone could cause the TESS signal.  Hypothetically, dust dedged up in the mass loss process could cause a large enough broadband flux loss to manifest in TESS.  We may expect to see some variable reddening in that case.  A putative H$\alpha$ line would have to be about 30\% deep over 10 \AA~ wide to manifest perceptibly in the TESS data.  Such an implausibly deep line likely would have been observed as perturbations in the Keck HIRES spectra, even with its limited phasing.

\subsection{Implications for exosphere detection in non-transiting planets}

We measured a large extent of Helium escape along the arc of the orbit, but we necessarily can place only coarse constraints on the vertical extent---in the direction perpendicular to the orbital path, $\pm z$.  We estimate the detectable vertical extent must be at least a few stellar radii, such that we still could have detected Helium escape in a hypothetical HAT-P-67b-like system even if it were non-transiting, in a ``near miss'' configuration.  We propose a new category of semi-transiting planet, dubbed ``exospheric grazers'', in which the planet does not produce a detectable white-light transit depth, but \emph{does} produce measureable \emph{line}-based exospheric absorption.  This category appears to have been neglected due to the assumption that exospheres were only easily detectable at several planetary radii, and therefore fractions of a stellar radius.  While large tails have been seen previously in Ly$\alpha$, the scarcity of UV resources prohibited searches of this kind.  The discovery of large Helium tails in HAT-P-32b (Zhang et al. 2023) and now HAT-P-67b suggest that the identification of these exospheric grazers may be achievable with current instrumentation and possibly existing archival data from near-IR RV-planet searches.

The configuration would consist of a HAT-P-67b-like planet undergoing profound mass loss, with a jet-like column of material either spouting from the poles, or lofted there through gas dynamics, or otherwise confined by bow-shock redirection.  A process for identifying strong candidates could select close-in planets or planet candidates discovered from non-transiting techniques, such as RVs or TTVs.  The best candidates would be planets with $M\sin{i}$, orbital periods, and insolations consistent with a hot Saturn mass.  A followup program could be devised to acquire in-transit spectroscopy to search for \ion{He}{1} 10830~\AA, or other spectral tracers of mass loss.

A few benefits of of this planet category arise.  First, they would expand the sample of planets for which the planetary winds could be measured, by relaxing the narrow geometrical transit probability of $1-10^\circ$ to something closer to $\pm30^\circ$, depending on the mass loss and gas dynamics.  In particular, the smallest stars such as white dwarfs and M-dwarfs could see the search space for exospheres increase by $10-100\times$.  Second, the range of viewing angles would serve as a natural experiment for constraining the vertical extent of outflows in the first place.  The typical bow shock 3D morphology could be deduced from a sample of otherwise-similar planets with a range of orbital inclinations.  Non-detections would therefore be valuable, especially in cases where $\sin{i}$ could be constrained independently.  In practice these additional constraints are rarely possible, so ensemble statistical methods may have to be employed.  Finally, for some promising planet candidates a distinctive helium signal could validate the planet, assuming the signal could be disambiguated from stellar variability.  That disambiguation may be feasible, since the velocity structure and phasing of stellar and planetary Helium signals diverge due to stellar wind confinement and the launching from a predictable planetary rest frame.  The elusive category of young planets could make ideal targets, since their large radii and high intrinsic temperatures and high incident XUV radiation may yield large mass loss rates if they have recently migrated to an interior location through disk migration.  This category is especially valuable if discovered through near-IR RV searches, since abundant multi-epoch spectroscopy may already contain the requisite stellar variability baselines.

\subsection{Interplay with rotating stellar disk} \label{vsini-interplay}

As aluded to previously, the interplay between the rotating stellar disk and the Helium line absorption sets up a subtle Doppler tomography-like signal.

We can understand this transmissive Doppler tomography effect through an example of four pathological limiting cases, illustrated in Figure \ref{fig:interplayvsini}.  Imagine the escaping atmsosphere were concentrated in a 20\% transmissive cylinder coming out of the orbital plane, as in an idealized transparent bipolar outflow.  This cyclinder would possess the planet's bulk velocity of about $-30$ km/s at the moment of transit ingress, placing the locus of absorption at about 10832.4 \AA.  But the absorptive cylinder would intercept a finite range of stellar limb velocities.  There would exist a constructive interference at the moment when the planet radial velocity shifts the \ion{He}{1} 10830 \AA line to share the same observed wavelength as the stellar lines at that limb position.

\begin{figure*}[t]
    \includegraphics[width=\linewidth]{figures/disperse_vsini_interplay_v0p1.pdf}
    \caption{Illustration of the interplay between a transiting columnar exosphere (extreme case) and a rotating stellar disk.  The top row shows a non-rotating (negligble $v\sin{i}$) case, with sharp stellar lines.  The motion of the planet can be seen easily, and the signals align at the planet's zero velocity (center of the disk).  The bottom row shows the rapidly rotating case.  Depending on the planet orbital speed and stellar rotational speeds, the signals can align at some pair wavelength before and after transit mid-center.  The alignment of the signals acts to \emph{suppress} the signal since the planet exosphere is blocking the fainter-on-average flux.  If the gas is spread out uniformly over the disk---instead of in a vertical column---the signals simply multiply.}
    \label{fig:interplayvsini}
\end{figure*}

\section{Conclusions}

\clearpage
\pagebreak

\appendix

\section{Spectral variability non-detections} \label{appendixSec}
\subsection{Calcium IR Triplet and other line diagnostics}
We examined the \ion{Ca}{2} IR Triplet lines (Ca IRT) for evidence of variability.  No variation was seen in these features, which were pre-processed in the same way as the Helium feature.  Figure \ref{fig:CaPhaseScan} shows a similar phase plot as Figure \ref{fig:HPFscanResid}, adapted to the line at 8662 \AA.  We see no conspicuous variability at this or any of the other Ca IRT lines.  We also found no detectable variability in the Pa$\delta$ line at 10050 \AA, nor other deep lines at 10330 \AA~ and elsewhere.  \ion{He}{1} 10833 appears to be the only conspicuously variable line in the HPF spectrum.

\begin{figure}[b]
    \includegraphics[width=0.5\linewidth]{figures/Ca8662_phase_2D_diagram_resid.png}
    \caption{Non-detection of variability in the Calcium IR Triplet line at 8862 $\AA$.  No other HPF lines appear to show any variability.}
    \label{fig:CaPhaseScan}
\end{figure}

\begin{acknowledgements}

    This material is based upon work supported by the National Aeronautics and Space Administration under Grant Number 80NSSC20K0257 for the XRP program issued through the Science Mission Directorate.

    Based on observations obtained with the Hobby-Eberly Telescope (HET), which is a joint project of the University of Texas at Austin, the Pennsylvania State University, Ludwig-Maximillians-Universitaet Muenchen, and Georg-August Universitaet Goettingen. The HET is named in honor of its principal benefactors, William P. Hobby and Robert E. Eberly.

    These results are based on observations obtained with the Habitable-zone Planet Finder Spectrograph on the HET. The HPF team acknowledges support from NSF grants AST-1006676, AST-1126413, AST-1310885, AST-1517592, AST-1310875, ATI 2009889, ATI-2009982, AST-2108512, and the NASA Astrobiology Institute (NNA09DA76A) in the pursuit of precision radial velocities in the NIR. The HPF team also acknowledges support from the Heising-Simons Foundation via grant 2017-0494.

    This paper includes data collected with the TESS mission, obtained from the MAST data archive at the Space Telescope Science Institute (STScI). Funding for the TESS mission is provided by the NASA Explorer Program. STScI is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5–26555.

    This research has made use of the Keck Observatory Archive (KOA), which is operated by the W. M. Keck Observatory and the NASA Exoplanet Science Institute (NExScI), under contract with the National Aeronautics and Space Administration.

    This research has made use of the NASA Exoplanet Archive, which is operated by the California Institute of Technology, under contract with the National Aeronautics and Space Administration under the Exoplanet Exploration Program.

    This research has made use of NASA's Astrophysics Data System.
\end{acknowledgements}

\facilities{HET (HPF), TESS, ASAS, Exoplanet Archive}

\software{  \texttt{pandas} \citep{mckinney10},
    \texttt{emcee} \citep{foreman13},
    \texttt{matplotlib} \citep{hunter07},
    \texttt{numpy} \citep{2020NumPy-Array},
    \texttt{scipy} \citep{2020SciPy-NMeth},
    \texttt{ipython} \citep{perez07},
    \texttt{seaborn} \citep{waskom14},
    \texttt{astropy} \citep{2022ApJ...935..167A},
    \texttt{muler} \citep{2022JOSS....7.4302G},
    \texttt{telfit} \citep{2014AJ....148...53G},
    \texttt{exoplanet} \citep{exoplanet:joss},
    \texttt{jupyter} \citep{Kluyver2016jupyter},
    \texttt{p-winds}, \citep{2022A&A...659A..62D}
}

\clearpage

\bibliography{ms}

\end{document}